---
title: "R Notebook"
output: html_notebook
---

Important packages:

```{r}
setwd("Dataset")
Dataset <- read.csv("data.csv")
Preprocessed_dataset <-  read.csv("preprocessed_dataset.csv") 
if(!require(ggplot2)){
install.packages("ggplot2")}
library(ggplot2)
if(!require(dplyr)){
install.packages("dplyr")}
library(dplyr) 
if(!require(dplyr)){
install.packages("dplyr")}
library(dplyr)
if(!require(ltm)){
install.packages("ltm")}
library(ltm)
if (!require(cluster, quietly = TRUE)) {
  install.packages("cluster")
}
if (!require(factoextra, quietly = TRUE)) {
  install.packages("factoextra")
}
 
```

# IT326 Project {style="color: gray"}

\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

## The Goal

Our primary objective of this analysis is to classify whatever a student will go to college or not using the classification methods and to identify the main factors and reasons why students are less likely to pursue higher education indicated by "will_go_to_college" being 'False'. By leveraging the provided dataset with attributes such as school type, school accreditation, gender, interest in college, residence, we aim to discover the most influential variables and their relationships with the decision not to attend college.

## The Source

Kaggle.com

## URL :

<https://www.kaggle.com/datasets/saddamazyazy/go-to-college-dataset>

## General information

-   Number of attributes : 11
-   Number of rows (objects) : 1000
-   The class label: The class label in our project is the attribute "will_go_to_college". This attribute is binary, meaning that it can take on two values: True for yes or False for no. The value of this attribute will be the target variable that we are trying to predict during our project.

------------------------------------------------------------------------

+-----------------------+------------------------------------------------------------+---------+---------------------------------------------------------------------------+
| Attribute             | Description                                                | Type    | Possible values                                                           |
+-----------------------+------------------------------------------------------------+---------+---------------------------------------------------------------------------+
| type_school           | The type of school the student attends                     | Binary  | Academic / Vocational                                                     |
+-----------------------+------------------------------------------------------------+---------+---------------------------------------------------------------------------+
| school_accreditation  | The quality if school                                      | Binary  | A / B (A is better than B)                                                |
+-----------------------+------------------------------------------------------------+---------+---------------------------------------------------------------------------+
| gender                | The student's gender                                       | Binary  | Male / Female                                                             |
+-----------------------+------------------------------------------------------------+---------+---------------------------------------------------------------------------+
| interest              | The student's interest in going to college                 | Nominal | Very interested /Interested / Less Interested / Not Interested /Uncertain |
+-----------------------+------------------------------------------------------------+---------+---------------------------------------------------------------------------+
| residence             | The student's residence                                    | Binary  | Urban / Rural                                                             |
+-----------------------+------------------------------------------------------------+---------+---------------------------------------------------------------------------+
| parent_age            | The age of the student's parents                           | Numeric | 40 - 65                                                                   |
+-----------------------+------------------------------------------------------------+---------+---------------------------------------------------------------------------+
| parent_salary         | The monthly salary of the student's parents in IDR/Rupiah. | Numeric | 1000K - 10M                                                               |
|                       |                                                            |         |                                                                           |
|                       | [1Rupiah = 0.00024SAR]                                     |         |                                                                           |
+-----------------------+------------------------------------------------------------+---------+---------------------------------------------------------------------------+
| house_area            | The size of the student's house in meter square            | Numeric | 20 - 120                                                                  |
+-----------------------+------------------------------------------------------------+---------+---------------------------------------------------------------------------+
| average_grades        | The student's average grades in school.                    | Numeric | 75 - 98                                                                   |
+-----------------------+------------------------------------------------------------+---------+---------------------------------------------------------------------------+
| parent_was_in_college | Whether the student's parents attended college.            | Binary  | True - False                                                              |
+-----------------------+------------------------------------------------------------+---------+---------------------------------------------------------------------------+
| will_go_to_college    | Whether the student plans to go to college.                | Binary  | True - False                                                              |
+-----------------------+------------------------------------------------------------+---------+---------------------------------------------------------------------------+

## Sample of our data

```{r}
head(Dataset)
```

Here are a sample of 6 row from our dataset.

## Missing values

```{r}
sum(is.na(Dataset))
```

There are no missing values in our dataset.

## Statistical graphs

Graph 1:

```{r}
df =data.frame(Dataset)
ggplot(data=df, aes(x = interest, fill = will_go_to_college)) +
  geom_bar() +
  scale_x_discrete(limits = c('Not Interested', 'Less Interested', 'Uncertain', 'Interested', 'Very Interested')) +
  labs(title = 'College interest vs College attendance ') +
  scale_fill_manual(values = c("True" = "antiquewhite2", "False" = "antiquewhite3")) +
  theme_minimal()


```

According to the graph, whether students are interested in going to college or not does not affect whether they actually end up attending Collage . There is a group of individuals who were interested in attending but did not receive acceptance, while others who were not interested were accepted.

Graph 2:

```{r}

filtered_True =filter(Dataset, will_go_to_college == 'True')
filtered_False =subset(Dataset, will_go_to_college =='False')

ggplot() +
  geom_density(data = filtered_True, aes(x = average_grades, fill = "Going to College"), alpha = 0.5) +
  geom_density(data = filtered_False, aes(x = average_grades, fill = "NOT Going to College"), alpha = 0.5) +
  labs(x = "Average Grades", y = "Density") +
  ggtitle("Comparison of Average Grades for Students Going to College and NOT Going to College") +
  scale_fill_manual(values = c("Going to College" = "antiquewhite4", "NOT Going to College" = "antiquewhite1"))

```

The graph shows that the average grades for students who had accepted to go to college were higher than those who did not enter college , and this indicates the existence of a correlation between those who going to college and the average grades

Graph 3:

```{r}
Dataset_percentage <- Dataset %>%
  group_by(type_school) %>%
  summarise(percentage = mean(will_go_to_college == "True") * 100)

# Create a percentage chart
ggplot(Dataset_percentage, aes(x = type_school, y = percentage, fill = type_school)) +
  geom_bar(stat = "identity") +
  labs(title = "Percentage of Students Going to College by Type of School",
       x = "Type of School",
       y = "Percentage") +
  scale_fill_manual(values = c("Academic" = "antiquewhite3", "Vocational" = "antiquewhite2")) +
  theme_minimal()

 
```

This graph shows the impact of the type of high school attended by students on their college attendance . Based on the bar chart:

-   among students from Academic high schools, 313 are going to college, and 296 are not.

-   among students from Vocational high schools, 187 are going to college, and 204 are not

These information tell us that a higher proportion of students from academic high schools are going to college compared to those from vocational schools which suggest that the type of school attended.

Graph 4:

```{r}
ggplot(Dataset, aes(x = average_grades)) +
  geom_histogram(binwidth = 5, fill = "antiquewhite2", color = "antiquewhite4") +
  labs(title = "Distribution of students' grades",
       x = "Students' average grades",
       y = "Frequency") +
  theme_minimal()
```

This histogram show us that the majority of the students in the dataset are performing well since it seems like their grades are spanning between 75 and 98. This analysis will help us determine whether the academic performance level of students is a contributing factor to their college attendance or not.

## Statistical Measures

1.  The student's academic performance analysis:

```{r}
summary(Dataset$average_grades)

```

The student grades in our dataset range from 75.00 to 98.00, with a median of 85.58 and an average of 86.10. This suggests that most students are doing well as none of them have average grades below 50. However, it's interesting to note that some students have much higher or lower grades than the average, mainly due to the wide range of grades..

2.  The socioeconomic status of students' families analysis:

```{r}
summary(Dataset$parent_salary)
```

In the dataset, we've got students parents with salaries ranging from 1,000,000 to 10,000,000 IDR/Rupiah. The median salary is 5,440,000 IDR/Rupiah, and the average is 5,381,570 IDR/Rupiah. This data tells us that many parents in our dataset earn less than the average salary in Indonesia, which is 146,000,000 IDR. This suggests that quite a few students in our dataset come from families with limited finances. And this financial situation could certainly impact their ability to get through college.

```{r}
summary(Dataset$house_area)
```

Additionally we can utilize the house area attribute to gain a deeper understanding of the socioeconomic status of students' families, where students with houses significantly larger than the mean might indicate a higher socioeconomic status, while those with houses considerably smaller than the mean might reflect a comparatively lower socioeconomic status. Based on the shown output, the house areas range from [20.00-120.00 ㎡]. The median house area is 75.50 ㎡ indicates that families with house areas around this value likely have moderate socioeconomic status with houses that neither very small nor very large.

3.  Understanding Parent Age Range and Variation in its Values

```{r}
summary(Dataset$parent_age)
SD=sd(Dataset$parent_age)
MeanAge=mean(Dataset$parent_age)
cat("coefficient of variation:",SD/MeanAge*100,"%")

```

This summary provides the range for age attribute [40,65] which indicates that all parent in middle age during this age parent have more concern about their children , the coefficient of variation= 6.7% which indicates lower variation ,and the value of attribute parent_agerare are relatively close to the mean overall 25% of them have an age below or equal to 50 , 75% have an age below or equal to 54 and the median value is 52

## Outliers analysis

```{r}
###parent age outliers
quartiles <- quantile(Dataset$parent_age, probs=c(.25, .75), na.rm = FALSE)
IQR <- IQR(Dataset$parent_age)

Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 

data_no_outlier <- subset(Dataset, Dataset$parent_age > Lower & Dataset$parent_age < Upper)
dim(data_no_outlier)

###parent salary outliers
quartiles <- quantile(Dataset$parent_salary, probs=c(.25, .75), na.rm = FALSE)
IQR <- IQR(Dataset$parent_salary)

Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 

data_no_outlier <- subset(data_no_outlier, data_no_outlier$parent_salary> Lower & data_no_outlier$parent_salary < Upper)
dim(data_no_outlier)


###averge grades outliers
quartiles <- quantile(Dataset$average_grades, probs=c(.25, .75), na.rm = FALSE)
IQR <- IQR(Dataset$average_grades)

Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 

data_no_outlier <- subset(data_no_outlier, data_no_outlier$average_grades> Lower & data_no_outlier$average_grades < Upper)
dim(data_no_outlier)


###house area outliers
quartiles <- quantile(Dataset$house_area, probs=c(.25, .75), na.rm = FALSE)
IQR <- IQR(Dataset$house_area)

Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 

data_no_outlier <- subset(data_no_outlier, data_no_outlier$house_area> Lower & data_no_outlier$house_area < Upper)

Founded_Outliers=data.frame(anti_join(Dataset,data_no_outlier))
print(Founded_Outliers)
```

After conducting data analysis and identifying outliers, our inspection reveals that the detected outliers represent inherent variation within the population. Regarding Parent_age, outliers are observed for values below 44 and above 65. However, it should be noted the age from 40 to 65 fall within the expected mean of our dataset meaning that it doesn't indicate that they are outliers . For parent_salary, we found two outliers: one below 1,326,250 ind ≈ 85 USD and another above 9,416,250 ind ≈ 606 USD. The minimum and maximum values were determined to be 1,000,000 ind ≈ 64 USD and 10,000,000 ind ≈ 644 USD, respectively. In the case of grades, twelve outliers were identified, ranging from below 76 to above 97. Nevertheless, since the data falls within the acceptable range of 0 to 100, these outliers should be retained as they are still considered normal and within the usual grade range. Finally, for house_area, we found eleven outliers below 34.4m and above 115m, with the minimum being 20m and the maximum being 120m. However, these values are still considered typical for the population.

## Normalization

```{r}
normalize <- function(x) {return((x-min(x))/ (max(x)-min(x)))}
datasetWithoutNormalization<-Dataset
Dataset$parent_salary<-normalize(datasetWithoutNormalization$parent_salary)
Dataset$house_area<-normalize(datasetWithoutNormalization$house_area)
print(Dataset)
```

We applied normalization to the 'parent_salary' and 'house_area' attributes, scaling their values to a range between 0 and 1. This normalization process greatly facilitates data handling and analysis, ensuring that these attributes are on a consistent scale. Which will improve the reliability of our data analysis and enable better conclusions to be drawn from the dataset. Normalization is a crucial step in preparing the data for modeling, as it prevents attributes with larger numerical ranges from dominating the analysis and ensures fair treatment for all features.

## Discretization

```{r}


Dataset$average_grades [Dataset$average_grades >= 95] <- '+A'
Dataset$average_grades [95 >Dataset$average_grades & Dataset$average_grades >= 90] <- 'A'
Dataset$average_grades [90 >Dataset$average_grades & Dataset$average_grades >= 85] <- '+B'
Dataset$average_grades [85 >Dataset$average_grades & Dataset$average_grades >= 80] <- 'B'
Dataset$average_grades [80 >Dataset$average_grades & Dataset$average_grades >= 75] <- '+C'
Dataset$average_grades [75 >Dataset$average_grades & Dataset$average_grades >= 70] <- 'C'
Dataset$average_grades [70 >Dataset$average_grades & Dataset$average_grades >= 65] <- '+D'
Dataset$average_grades [65 >Dataset$average_grades & Dataset$average_grades >= 60] <- 'D'
Dataset$average_grades [60 >Dataset$average_grades & Dataset$average_grades >= 0] <- 'F'
Dataset$average_grades <- as.character(Dataset$average_grades )
print(Dataset)
```

We transformed the parent_age attribute into intervals by dividing the values to be fall on one of two possible interval labels with equal width which is(40,50],(50,60] by discretization the values well be simpler to classify or perform other methods that can help us later in our model.

and to better utilize and interpret the grades attributes for each student, we have converted the numeric grades into letter grades (A+, A, B+, B, C+, C, D+, D, F). This transformation was undertaken to focus on the general letter grade representation rather than the precise numerical values.

## Encoding

```{r}
Dataset$parent_was_in_college[Dataset$parent_was_in_college=="TRUE"]<-1
Dataset$parent_was_in_college[Dataset$parent_was_in_college=="True"]<-1
Dataset$parent_was_in_college[Dataset$parent_was_in_college=="FALSE"]<-0
Dataset$parent_was_in_college[Dataset$parent_was_in_college=="False"]<-0

Dataset$will_go_to_college[Dataset$will_go_to_college=="TRUE"]<-0
Dataset$will_go_to_college[Dataset$will_go_to_college=="True"]<-0
Dataset$will_go_to_college[Dataset$will_go_to_college=="FALSE"]<-1
Dataset$will_go_to_college[Dataset$will_go_to_college=="False"]<-1

Dataset$gender[Dataset$gender=="Female"]<-1
Dataset$gender[Dataset$gender=="Male"]<-0

Dataset$school_accreditation[Dataset$school_accreditation=="A"]<-1
Dataset$school_accreditation[Dataset$school_accreditation=="B"]<-0

Dataset$interest[Dataset$interest=="Very Interested"]<-4
Dataset$interest[Dataset$interest=="Interested"]<-3
Dataset$interest[Dataset$interest=="Less Interested"]<-2
Dataset$interest[Dataset$interest=="Not Interested"]<-1
Dataset$interest[Dataset$interest=="Uncertain"]<-0


Dataset$type_school[Dataset$type_school=="Academic"]<-1 
Dataset$type_school[Dataset$type_school=="Vocational"]<-0

Dataset$residence[Dataset$residence=="Urban"]<-1
Dataset$residence[Dataset$residence=="Rural"]<-0
print(Dataset)
```

Since encoding is an important step in data preprocessing that enables the use of categorical data in various data analysis and machine learning tasks, we encoded attributes like the 'parent was in college' attribute from (True, False) to (1, 0), and 'will go to college' from (True, False) to (0, 1). This encoding is carried out as we aim to predict the influencing factors. Additionally, we encoded the 'gender' attribute from (Female, Male) to (1, 0), 'school accreditation' from (A, B) to (1, 0), 'type_school' from (Academic, Vocational) to (1, 0), 'residence' from (Urban, Rural) to (1, 0), and 'interest' from (Very interested ,Interested , Less Interested , Not Interested ,Uncertain ) to (4,3,2, 1, 0) respectively. Encoding serves to simplify the data, reduce complexity, and enhance its suitability for modeling purposes.

## Correlation analysis Chi square test for nominal attribute:

```{r}

#1
C=chisq.test(Dataset$type_school , Dataset$will_go_to_college)
print(C)
#2

C=chisq.test(Dataset$school_accreditation , Dataset$will_go_to_college)
print(C)
#3
C=chisq.test(Dataset$gender , Dataset$will_go_to_college)
print(C)
#4
C=chisq.test(Dataset$interest , Dataset$will_go_to_college)
print(C)

#5
C=chisq.test(Dataset$residence , Dataset$will_go_to_college)
print(C)

#6
C=chisq.test(Dataset$average_grades , Dataset$will_go_to_college)
print(C)

#7
C=chisq.test(Dataset$parent_was_in_college , Dataset$will_go_to_college)
print(C)
```

All the attributes have X-square greater than the p-value which indicate a some association with the class label; therefore we reject the null hypothesis

we noticed for 'interest' and 'average grade' the analysis shows that X-square is much larger than p-value indicate the significant association of the two attributes with the decision of the student to go to the collage or not

## Correlation coefficient analysis for numeric attribute:

```{r}

biserial.cor(Dataset$parent_salary,Dataset$will_go_to_college, c("all.obs", "complete.obs"), level = 1)
biserial.cor(Dataset$house_area,Dataset$will_go_to_college, c("all.obs", "complete.obs"), level = 1)
biserial.cor(Dataset$parent_age,Dataset$will_go_to_college,c("all.obs", "complete.obs"), level = 1)
 
 

```

the analysis shows moderate correlation coefficient for parent salary and house area with the class label which indicate that they are relevant factors meaning that the higher the parent salary and the larger house area the higher probability for a student to enroll in a collage

where is the on other hand, the correlation coefficient for the parent age is very small which indicate that the parent age has little impact to the probability for student to enroll in a collage

## Feature selection:

ultimately based on the analysis of the correlation that we conducted on the relationship of the dataset attributes with the class label, and the understanding of the data and the context of each attribute and potential relevance to the class label we decided to not delete any of the attribute



